---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# powRICLPM

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/powRICLPM)](https://CRAN.R-project.org/package=powRICLPM)
<!-- badges: end -->

`powRICLPM` is an `R` package that performs a power analysis for the random intercept cross-lagged panel model (RI-CLPM) in a simple and use-friendly way. Its main functionalities are:

1. [Recommend a sample size](https://jeroendmulder.github.io/powRICLPM/articles/recommend.html) for a desired power level for a specific parameter.
2. [Obtain performance measures](https://jeroendmulder.github.io/powRICLPM/articles/performance.html) (e.g., bias, mean square error, etc.) for parameters across simulation conditions with varying sample size, number of time points, and proportion of between-unit variance.
3. [Create Mplus syntax](https://jeroendmulder.github.io/powRICLPM/articles/mplus.html) for performing a power analysis for the RI-CLPM using Mplus. 

User guides for the above functionalities can be found in vignettes under the 'Articles' tab. Technical details on the implementation of the power analysis can be found in LINK NAAR PAPER. 

## Installation

You can install the development version of `powRICLPM` from GitHub with:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("jeroendmulder/powRICLPM")
```

## Documentation

Every user-facing function in the package is documented, and the documentation can be accessed by running `?function_name` in the R console, e.g., `?powRICLPM`. Furthermore, there are four main vignettes (accessible via the 'Articles' tab), describing functionalities and analysis options of this package.

## Citing `powRICLPM`

You can cite the R-package with the following citation:

> Mulder, J.D., (n.d.). *Performing power analysis for the RI-CLPM* 

## Contact

If you have ideas, comments, or issues you would like to raise, please get in touch.  

* Issues and idea can be raised on GitHub via [https://github.com/jeroendmulder/powRICLPM](https://github.com/jeroendmulder/powRICLPM)
* Pull request can be raised on GitHub via [https://github.com/jeroendmulder/powRICLPM/pulls](https://github.com/jeroendmulder/powRICLPM/pulls)

